---
  title: "Descriptive Analysis, PCA and Factor Model"
author: "Barbara Bettencourt"
date: "26/11/2021"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Stocks:**
  Airlines:  
  Lufthansa (German); DLAKY    
Southwest Airlines; LUV   
Easy Jet (British); EZJ    
RyanAir Holdings plc (British); RYAAY    
United Airlines Holdings Inc (American); UAL    
American Airlines Group Inc (American); AAL    
Spirit Airlines Inc (American); SAVE  

Beer Industry:  
  Carlsberg (Danish); CABGY
Heineken (Netherlands); HEINY  
Anheuser-Busch InBev (Belgium); BUD  
Harboe (Danish); CPH: HARB-B  
Guinness from Diageo; DEO  
Molson Coors Brewing; TAP
Tsingtao Brewery Group; OTCMKTS: TSGTF
Asahi Group Holding Ltd; OTCMKTS: ASBRF

**Descriptive Analysis**

```{r include=FALSE}
#libraries

library(tidyquant)
library(purrr)
library(dplyr)
library(tidyverse)
library(Hmisc)
library(reshape)
library(tseries)
library(quadprog)
library(DescTools)
library(ggplot2)
library(reshape2)
library(car)
library(tseries)
library(gridExtra)
library(RCurl)
library(readr)
library(QRM)
library("Ecdat")
library("fGarch")

```

This is how we collected the data originally, however, somehow some of the Asahi returns disappeared after I've written most of this code, so we use the data.csv file (the original data). The returns changed on the 10th of December.

```{r include=FALSE}

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers <- c(
  'DLAKY',
  'LUV',
  'EZJ',
  'RYAAY',
  'UAL',
  'AAL',
  'SAVE',
  'CABGY',
  'HEINY',
  'BUD',
  'HARB-B.CO',
  'DEO',
  'TAP',
  'TSGTF',
  'ASBRF',
  'IVV'
)

stock_names <- c(
  'Lufthansa',
  'Southwest',
  'EasyJet',
  'RyanAir',
  'United',
  'American',
  'Spirit',
  'Carlsberg',
  'Heineken',
  'AB',
  'Harboe',
  'Guinness',
  'Molson',
  'Tsingtao',
  'Asahi',
  'SP500'
)

stocks <- data.frame(tickers = tickers,
                     stockname = stock_names)

```

```{r}

returns <- tq_get(tickers, get="stock.prices") %>%
  group_by(symbol) %>%
  tq_transmute(select=adjusted,
               mutate_fun=periodReturn,
               period="monthly",
               col_rename = "monthly_return")

returns <- returns %>% 
  pivot_wider(names_from = symbol, values_from = monthly_return)

#deleted a bunch of NA Values
returns <- returns[1:131,]

```

```{r}
#here i replace the entire data that was collected online with the data.csv file as it is the only way my code would work again after the sudden change in the online returns

returns3 <- read.csv('https://raw.githubusercontent.com/barbarabettencourt/Project-Finance/main/data/data.csv')
returns3 <- returns3[2:17]

returns[,2:16] <- returns3[,2:16]
```

```{r include=FALSE}
rm(returns3);rm(tickers)
```


**Descriptive Analysis**
1. Table of mean, sd, skewness, kurtosis and betas.   

```{r}

returns <- as.data.frame(returns)
returnsNA <- as.data.frame(na.omit(returns))

# columns of table
Stocks <- stock_names[1:15]
Means <- rep(NA, 15)
Sd <- rep(NA,15)
Skewness <- rep(NA,15)
Kurtosis <- rep(NA,15)
Betas <- rep(NA,15)
table <- data.frame(Stocks,Means,Sd,Skewness,Kurtosis,Betas)

# means

for(i in 1:15)
{
  table$Means[i] <- mean(returnsNA[,i+1])
}

# sd

for(i in 1:15)
{
  table$Sd[i] <- sd(returnsNA[,i+1])
}

# skewness should be around 0 to be normally distributed

for(i in 1:15)
{
  table$Skewness[i] <- Skew(returnsNA[,i+1])
}

# kurtosis should be around 3 to be normally distributed

for(i in 1:15)
{
  table$Kurtosis[i] <- Kurt(returnsNA[,i+1])
}

# betas using riskfree rate of 0.45 with SML

risk.free <- rep(0.45,113)

regressions <- vector("list",15)

for (i in 1:15) 
{
  regressions[[i]] <- lm((returnsNA[,i+1]-risk.free)~(returnsNA[,17]-risk.free))  
}


for(i in 1:15)
{
  table$Betas[i] <- regressions[[i]][[1]][2]
}

table

```
```{r include=FALSE}
rm(Betas);rm(regressions);rm(i);rm(Kurtosis);rm(Means);rm(risk.free);rm(Sd);rm(Skewness)
```


2. Plot each set of returns  

```{r}

# Individual stocks

for (i in 1:15)
{
  names <- list(stock_names)
  print(ggplot(returns, aes(x=date, y=returns[,i+1])) +
          geom_line(color="steelblue") + 
          geom_point() +
          xlab("") +
          ylab(names[[1]][i])+
          ggtitle("Plot of Returns") + 
          theme(plot.title = element_text(hjust = 0.5))) 
}

# United airlines and American airlines

airlines <- returns %>%
  select(date, UAL, AAL) %>%
  gather(key="Stocks", value = "Returns", -date)

ggplot(airlines, aes(x = date, y = Returns)) + 
  geom_line(aes(color = Stocks)) + 
  scale_color_manual(values = c("skyblue4", "orchid1")) +
  geom_hline(yintercept=0,linetype="dashed", alpha=0.5)+
  geom_vline(xintercept = as.Date("2020-01-31"), linetype="dashed", alpha=0.5) +
  ggtitle("Returns of United and American Airlines") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("")

rm(airlines)

# Carslberg and Asahi

beers <- returns %>%
  select(date, CABGY, ASBRF ) %>%
  gather(key="Stocks", value = "Returns", -date)

ggplot(beers, aes(x = date, y = Returns, group=1)) + 
  geom_line(aes(color = Stocks)) + 
  scale_color_manual(values = c("skyblue4", "orchid1")) +
  geom_hline(yintercept=0,linetype="dashed", alpha=0.5)+
  geom_vline(xintercept = as.Date("2020-01-31"), linetype="dashed", alpha=0.5) +
  ggtitle("Returns of Carlsberg and Asahi") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("")


rm(beers)
```
3. Plot equity curve

```{r}

returnsNA <- as.data.frame(na.omit(returns))   #doesn't work with NA values

equityCurve <- as.data.frame(matrix(c(rep(NA,113*15)), nrow= 113))
for (i in 1:15) 
{  
  equityCurve[,i] <- cumprod(1 + returnsNA[,i+1]) 
}

equityCurve <- cbind(returnsNA[,1], equityCurve) #adding date to dataframe
colnames(equityCurve)[1] <- "date"

#airlines individual
for (i in 1:7)
{
  names <- list(stock_names)
  print(ggplot(equityCurve, aes(x=date, y=equityCurve[,i+1])) +
          ggtitle("Equity Curve") +
          theme(plot.title = element_text(hjust = 0.5))+  
          geom_line(color="steelblue") + 
          xlab("") +
          ylab(names[[1]][i]))  
}

#beers individual
for (i in 1:8)
{
  names <- list(stock_names)
  print(ggplot(equityCurve, aes(x=date, y=equityCurve[,i+7])) +
          ggtitle("Equity Curve") + 
          theme(plot.title = element_text(hjust = 0.5)) +  
          geom_line(color="steelblue") + 
          xlab("") +
          ylab(names[[1]][i+7]))  
}

```
3.2 You should also provide an equity curve for each asset (that is, a curve that shows the growth of a $1 in each of the asset over the time period you chose) and comment of your results. You should do the same for S&P 500 and compare it with the assets.

```{r}

equityCurveSP <- cumprod(1 + returnsNA[,17]) #equity curve for SP500

namesair <- names[[1]][1:7]
namesair <- append(namesair, "SP500")

colors <- c(1,2,3,4,5,6,7,1)
lt <- c(1,1,1,1,1,1,1,2)

#airlines together with SP
plot(equityCurve[,1], equityCurve[,2], type = "l", ylim= c(0.2,6), xlab="", ylab="Airlines", main="Equity Curves of Airlines")
for(i in 1:6)
{
  lines(equityCurve[,1], equityCurve[,i+2], type = "l", col = i+1)
}
lines(equityCurve[,1],equityCurveSP,type="l", lty=2, lwd=1, col="black")
abline(v=as.Date("2020-01-31"), lty=2, col="dimgrey")
legend(x='topleft', legend=namesair,col=colors, lty=lt, cex=0.8)

####################################################################

colors <- c(1,2,3,4,5,6,7,8,1)
lt <- c(1,1,1,1,1,1,1,1,2)

#beers together with SP
plot(equityCurve[,1], equityCurve[,9], type = "l", ylim= c(0.4,4), xlab="", ylab="Beers", main="Equity Curves of Beers")
for(i in 1:7)
{
  lines(equityCurve[,1], equityCurve[,i+9], type = "l", col = i+1)
}
lines(equityCurve[,1],equityCurveSP,type="l", lty=2, lwd=1, col="black")
abline(v=as.Date("2020-01-31"), lty=2, col="dimgrey")
legend(x='topleft', legend=names[[1]][8:16],col=colors, lty=lt, cex=0.8)

```
```{r}
rm(equityCurve);rm(colors);rm(equityCurveSP);rm(lt);rm(namesair)
```

4. Histograms for each return series  

```{r}

for (i in 1:15)
{
  names <- list(stock_names)
  print(ggplot(returnsNA, aes(x=returnsNA[,i+1])) +
          geom_histogram(aes(y=..density..), colour="black", fill="white") + 
          geom_density(alpha=.2, fill="#FF6666")+ 
          xlab("") +
          ggtitle(names[[1]][i])+
          theme(plot.title = element_text(hjust = 0.5)))  
}

```
5. Boxplots for each return series  

```{r}

#individual boxplots

for (i in 1:15)
{
  names <- list(stock_names)
  boxplot(returns[,i+1], xlab="", main = names[[1]][i])
}

returnsbox <- returnsNA
colnames(returnsbox)[2:16] <- names[[1]]
colors <- c('cadetblue1', 'coral', 'darkolivegreen1', 'lightgoldenrod1')

#airlines
boxplot(returnsbox[,2:8], col = colors, main= "Boxplots of Airlines")

#beers
boxplot(returnsbox[,9:16], col= colors, main = "Boxplots of Beers")

```
```{r include=FALSE}
rm(returnsbox);rm(colors)
```


6. qqplots for each return series  

```{r}

for (i in 1:15)
{
  qqPlot(returnsNA[,i+1], main = names[[1]][i])
}

```

7. Run tests for stationarity  

```{r}

for (i in 1:15)
{
  print(adf.test(returnsNA[,i+1]))
}

#all pvalues are less than 0.01 and so p-value<0.05 hence we reject the null hypothesis and conclude that there is enough evidence to conclude that the returns are stationary

```

8. Check if returns are normally distributed

```{r}

for (i in 1:15)
{
  print(shapiro.test(returnsNA[,i+1]))
}

#results

#normal stocks <- 1,2,3,4,9,10,12,14,8 barely with 0.05991
#not normal stocks <- 5,6,7,11,13,15

```

9. Fit different distributions to your data, which one fits better?
  
```{r}

#t distribution

AIC_t <- c(rep(NA), 15)
BIC_t <- c(rep(NA), 15)

for (i in 1:15) 
{
  x <- diff(returnsNA[,i+1])
  n = length(x)
  start = c(mean(x), sd(x), 5)
  loglik_t = function(beta) sum( - dt((x - beta[1]) / beta[2], beta[3], log = TRUE) + log(beta[2]) )
  fit_t = optim(start, loglik_t, hessian = T, method = "L-BFGS-B", lower = c(-1, 0.001, 1))
  AIC_t[i] = 2 * fit_t$value + 2 * 3
  BIC_t[i] = 2 * fit_t$value + log(n) * 3  
}

#skewed student t
#doesnt work for Harboe (11) 

AIC_sstd <- c(rep(NA), 15)
BIC_sstd <- c(rep(NA), 15)

for (i in 1:10)
{  
  x <- diff(returnsNA[,i+1])
  n = length(x)
  loglik_sstd = function(beta) sum(- dsstd(x, mean = beta[1], sd = beta[2], nu = beta[3], xi = beta[4], log = TRUE))
  start = c(mean(x), sd(x), 5, 1)
  fit_sstd = optim(start, loglik_sstd, hessian = T, method = "L-BFGS-B", lower = c(-0.1, 0.01, 2.1, -2))
  AIC_sstd[i] = 2*fit_sstd$value + 2 * 4
  BIC_sstd[i]= 2*fit_sstd$value + log(n) * 4
}

for (i in 1:4)
{  
  x <- diff(returnsNA[,i+12])
  n = length(x)
  loglik_sstd = function(beta) sum(- dsstd(x, mean = beta[1], sd = beta[2], nu = beta[3], xi = beta[4], log = TRUE))
  start = c(mean(x), sd(x), 5, 1)
  fit_sstd = optim(start, loglik_sstd, hessian = T, method = "L-BFGS-B", lower = c(-0.1, 0.01, 2.1, -2))
  AIC_sstd[i+11] = 2*fit_sstd$value + 2 * 4
  BIC_sstd[i+11]= 2*fit_sstd$value + log(n) * 4
}

#ged 

AIC_ged <- c(rep(NA), 15)
BIC_ged <- c(rep(NA), 15)

for (i in 1:15)
{
  x <- diff(returnsNA[,i+1])
  n = length(x)
  loglik_ged = function(beta) sum( - dged(x,mean=beta[1],sd=beta[2],nu=beta[3],log=TRUE) )
  start = c(mean(x),sd(x),1)
  fit_ged = optim(start,loglik_ged,hessian=T,method="L-BFGS-B",lower = c(-.1,.01,1))
  AIC_ged[i] = 2*fit_ged$value+2*4
  BIC_ged[i] = 2*fit_ged$value+log(n)*4
}

#normal 
AIC_n <- c(rep(NA), 15)
BIC_n <- c(rep(NA), 15)

for (i in 1:15)
{
  x <- diff(returnsNA[,i+1])
  n = length(x)
  start = c(mean(x),sd(x),1)
  loglik_n = function(beta) sum( - dnorm(x,mean=beta[1],sd=beta[2],log=TRUE) )
  fit_n = optim(start, loglik_n, hessian = T,
                method = "L-BFGS-B", lower = c(-1, 0.001, 1))
  AIC_n[i] = 2 * fit_n$value + 2 * 3
  BIC_n[i] = 2 * fit_n$value + log(n) * 3
}

################################################################################

table2 <- data.frame(AIC_t,AIC_sstd,AIC_ged,AIC_n,BIC_t,BIC_sstd,BIC_ged,BIC_n)
table2 <- cbind(Stocks, table2)

AIC_result <- c("Skewed Sd t", "t Dist", "Ged", "Skewed Sd t", "t Dist", "t Dist", "t Dist", "Skewed Sd t", "t Dist", "t Dist", "Normal", "Skewed Sd t", "Skewed Sd t", "Normal", "t Dist")
BIC_result <- c("t Dist","t Dist","t Dist","t Dist","t Dist","t Dist","t Dist","t Dist","t Dist","t Dist","Normal","t Dist","t Dist","Normal","t Dist")

table2 <- cbind(table2, AIC_result, BIC_result)
colnames(table2)[2:11] <- c("t AIC", "Skewed t AIC", "Ged AIC", "Normal AIC", "t BIC", "Skewed t BIC", "Ged BIC", "Normal BIC", "Best AIC", "Best BIC")

table2
```
```{r include=FALSE}
rm(AIC_ged);rm(AIC_n);rm(AIC_result);rm(AIC_sstd);rm(AIC_t);rm(BIC_ged);rm(BIC_n);rm(BIC_result);rm(BIC_sstd);rm(BIC_t);rm(fit_ged);rm(fit_n);rm(fit_sstd);rm(fit_t);rm(table2);rm(n);rm(start);rm(x);rm(loglik_ged);rm(loglik_n);rm(loglik_sstd);rm(loglik_t);rm(Stocks)
```


10.  Compute Sharpe’s slope for each asset. Which asset has the highest slope?   
If E(RP) and $\sigma_{R_p}$ are the expected return and standard deviation of the return on a portfolio and $\mu_f$ is the risk-free rate, then
Sharpe's slope = $\frac{(E(R_P)-\mu_f)}{\sigma_{Rp}}$

```{r}
#risk free given by Professor = 0.45

risk.free <- 0.45

SharpesSlope <- rep(NA,15)

for (i in 1:15) 
{
  SharpesSlope[i] <- (mean(returnsNA[,i+1]) - risk.free)/sd(returnsNA[,i+1])
}

SharpesSlope

#If the results have all negative Sharpe ratio, it either means the risk-free rate is greater than the portfolio's return (which is the case here), or the portfolio's return is expected to be negative. In either case, a negative Sharpe ratio does not convey any useful meaning.

max(SharpesSlope) #highest slope

```
```{r include=FALSE}
rm(SharpesSlope)
```


11. Convert the monthly sample means into annual estimates by multiplying by 12 and convert the monthly sample SDs into annual estimates by multiplying by the square root of 12. Comment on the values of these annual
numbers.

```{r}
#annual means

annualsamplesmeans <- rep(NA,15)

for (i in 1:15) 
{
  annualsamplesmeans[i] <- mean(returnsNA[,i+1])*12

}

diffmeans <- cbind(annualsamplesmeans, table$Means)
diffmeans

#annual sd

annualsamplesd <- rep(NA,15)

for (i in 1:15) 
{
  annualsamplesd[i] <- sd(returnsNA[,i+1])*sqrt(12)

}

diffsd <- cbind(annualsamplesd, table$Sd)
diffsd


```
```{r include=FALSE}
rm(annualsamplesd);rm(annualsamplesmeans);rm(diffmeans);rm(diffsd)
```


12. Construct pairwise scatter plots between your assets returns and comment on
any relationships you see.

```{r}
#airlines
pairs(returns[,2:8], main="Scatterplot of Airline Returns")

#beers
pairs(returns[,9:16], main="Scatterplot of Beer Returns")
```

13. You should also compute the sample covariance matrix of the
returns on your assets and comment on the direction of linear association between the asset
returns.

```{r}

# covariance monthly data

Covariances <- as.data.frame(cov(returnsNA[2:16]))

#they all have positive covariances hence the returns all move together, vary in the same direction.

Covariancesair <- Covariances[1:7,1:7]
Covariancesbeers <- Covariances[8:15,8:15]

# covariance annualized data

annualreturns <- returnsNA[,2:16]*12

Covariancesannual <- as.data.frame(cov(annualreturns[1:15]))

#they all have positive covariances hence the returns all move together, vary in the same direction.

Covariancesairannual <- Covariancesannual[1:7,1:7]
Covariancesbeersannual <- Covariancesannual[8:15,8:15]

###############################################################################

#correlation for monthly data

Correlations <- as.data.frame(cor(returnsNA[2:16]))

Correlationsair <- Correlations[1:7,1:7]
Correlationsbeers <- Correlations[8:15,8:15]

```

```{r include=FALSE}
rm(Correlations);rm(Correlationsair);rm(Correlationsbeers);rm(Covariances);rm(Covariancesair);rm(Covariancesbeers);rm(Covariancesannual);rm(annualreturns);rm(Covariancesbeersannual);rm(Covariancesairannual)

```


**PCA**

```{r include=FALSE}
library(ggfortify)
```


```{r}
#all stocks

pca = prcomp(returnsNA[,2:16], scale = TRUE)

#pc1 and pc2
autoplot(pca, data=returnsNA[,2:16], loadings=TRUE, loadings.label=TRUE) + ggtitle("1st and 2nd Principal Components of the 15 stocks") + theme(plot.title = element_text(hjust = 0.5))

#pc2 and pc3
autoplot(pca, data=returnsNA[,2:16], loadings=TRUE, loadings.label=TRUE, x=2, y=3) + ggtitle("2nd and 3rd Principal Components of the 15 stocks") + theme(plot.title = element_text(hjust = 0.5))

summary(pca)

#airlines

pca = prcomp(returnsNA[,2:8], scale = TRUE)

#pc1 and pc2
autoplot(pca, data=returnsNA[,2:8], loadings=TRUE, loadings.label=TRUE)+ ggtitle("1st and 2nd Principal Components of the Airline stocks") + theme(plot.title = element_text(hjust = 0.5))

#pc2 and pc3
autoplot(pca, data=returnsNA[,2:8], loadings=TRUE, loadings.label=TRUE, x=2, y=3)+ ggtitle("2nd and 3rd Principal Components of the Airline stocks") + theme(plot.title = element_text(hjust = 0.5))

summary(pca)

#beers

pca = prcomp(returnsNA[,9:16], scale = TRUE)

#pc1 and pc2
autoplot(pca, data=returnsNA[,9:16], loadings=TRUE, loadings.label=TRUE)+ ggtitle("1st and 2nd Principal Components of the Beer stocks") + theme(plot.title = element_text(hjust = 0.5))

#pc2 and pc3
autoplot(pca, data=returnsNA[,9:16], loadings=TRUE, loadings.label=TRUE, x=2, y=3)+ggtitle("2nd and 3rd Principal Components of the Beer stocks") + theme(plot.title = element_text(hjust = 0.5))

summary(pca)
```
```{r include=FALSE}
rm(pca)
```


**Factor Models**

```{r}

fit <- factanal(x=returnsNA[2:16], factors = 2)
fit

#loadings are beta.hat transpose
#uniquenesses: the diagonal elements of the estimate bigsigma.hat_epsilon

#when we have two factors it appears that the airlines were grouped into factor 1 and the beers grouped in factor 2.

```