---
title: "Finance"
author: "Barbara, Frederik, Sierra"
date: "26/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Stocks:**
Airlines:  
Lufthansa (German); DLAKY    
Scandinavian Airlines (Danish); SAS    
Norwegian Air Shuttle (Norway); NWARF    
Easy Jet (British); EZJ    
RyanAir Holdings plc (British); RYAAY    
United Airlines Holdings Inc (American); UAL    
American Airlines Group Inc (American); AAL    
Swiss Air (American) (coundnt find) 
Spirit Airlines Inc (American); SAVE  

Beer Industry:  
Carlsberg (Danish); CABGY
Heineken (Netherlands); HEINY  
Anheuser-Busch InBev (Belgium); BUD  
Harboe (Danish); CPH: HARB-B  
Guinness from Diageo; DEO  
Molson Coors Brewing; TAP
Tsingtao Brewery Group; OTCMKTS: TSGTF
Asahi Group Holding Ltd; OTCMKTS: ASBRF

```{r include=FALSE}
library(tidyquant)
library(purrr)
library(dplyr)
library(tidyverse)

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)
```

```{r}

#example for stock of Lufthansa

time <- today()-years(5)
stocks.Lufthansa <- tq_get("DLAKY", get="stock.prices", from=time)

#daily returns

daily.Lufthansa <- stocks.Lufthansa %>% tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "daily", col_rename = "daily_returns")

head(daily.Lufthansa)

#monthly returns

monthly.Lufthansa <- stocks.Lufthansa %>%
  tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "monthly", col_rename = "monthly_returns")

head(monthly.Lufthansa)

```

