## Risk Management
100,000
VaR(0.05)
ES
over one month
normal dist
estimated means and vars
do nonparametric additionally
comment: highest lowest VaR and ES assets
bootstrap to estimate se and 95% CI for VaR(0.05) and ES

```{r}
# read data
data = read.csv('data.csv')
dim(data)
head(data)
stock_names <- c(
  'Lufthansa',
  'Southwest',
  'EasyJet',
  'RyanAir',
  'United',
  'American',
  'Spirit',
  'Carlsberg',
  'Heineken',
  'AB',
  'Harboe',
  'Guinness',
  'Molson',
  'Tsingtao',
  'Asahi'
)
```

```{r}
# read ret_p
ret_p = read.csv('')
dim(ret_p)
head(ret_p)
```


###### Param VaR, ES
```{r}
# param VaR and ES functions on normal distribution
VaR_norm = function(s0, mu, sigma, alpha) {
    return( -s0 + sigma * qnorm(alpha) )
}

ES_norm = function(s0, mu, sigma, alpha) {
    return( -s0 * ( -mu + sigma * ( pnorm( qnorm(alpha) ) / alpha ) ) )
}
```

```{r}
# estimate mean and sd
mean_vec = rep(NA, dim(data)[2] - 2)
sd_vec = rep(NA, dim(data)[2] - 2)
for (i in 1:(dim(data)[2] - 2)) {
    mean_vec[i] = mean(na.omit(data[, i + 2]))
    sd_vec[i] = sd(na.omit(data[, i + 2]))
}
```

```{r}
# param VaR, ES for assets
VaR_norm_ls = list()
ES_norm_ls = list()
for (i in 1:(dim(data)[2] - 2)) {
    name = stock_names[i]
    VaR_norm_ls[[name]] = VaR_norm(100000, mean_vec[i], sd_vec[i], 0.05)
    ES_norm_ls[[name]] = ES_norm(100000, mean_vec[i], sd_vec[i], 0.05)
}
```

```{r}
# estimate mean and sd for returns of portfolios
mean_p_vec = rep(NA, dim(ret_p))
sd_p_vec = rep(NA, dim(ret_p))
for (i in 1:dim(ret_p)) {
    mean_p_vec[i] = mean(ret_p[, i])
    sd_p_vec[i] = sd(ret_p[, i])
}
```

```{r}
# param VaR, ES for portfolios
VaR_norm_p_ls = list()
ES_norm_p_ls = list()
for (i in 1:dim(ret_p)) {
    name = names(ret_p)
    VaR_norm_p_ls[[name]] = VaR_norm(100000, mean_p_vec[i], sd_p_vec[i], 0.05)
    ES_norm_p_ls[[name]] = ES_norm(100000, mean_p_vec[i], sd_p_vec[i], 0.05)
}
```

###### Nonparam VaR, ES
```{r}
# nonparametric VaR and ER
VaR <- function(s0, ret, alpha) {
  return(-s0 * quantile(ret, alpha))
}

ES <- function(s0, ret, alpha) {
  return(-s0 * sum(ret[ret < -quantile(ret, alpha)] / length(ret[ret < -quantile(ret, alpha)])))
}
```

```{r}
# nonparam VaR, ES for assets
VaR_ls = list()
ES_ls = list()
for (i in 1:(dim(data)[2] - 2)) {
    name = stock_names[i]
    VaR_ls[[name]] = VaR(100000, na.omit(data[, i + 2]), 0.05)
    ES_ls[[name]] = ES(100000, na.omit(data[, i + 2]), 0.05)
}
```

```{r}
# nonparam VaR, ES for returns of portfolios
VaR_p_ls = list()
ES_p_ls = list()
for (i in 1:dim(ret_p)) {
    name = names(ret_p)
    VaR_p_ls[[name]] = VaR(100000, data[, i + 2], 0.05)
    ES_ls[[name]] = ES(100000, data[, i + 2], 0.05)
}
```

Highest VaR: Southwest, EasyJet, RyanAir, Carlsberg, Heineken, AB, Harboe, Guinness, Molson, Asahi
Lowest VaR: Tsingtao, Spirit, American, United, Lufthansa

###### Boostrap for se, 95% CI for VaR, ES 

```{r}
# se and 95% CI for VaR, ES of assets using boostrap
VaR_se = list()
ES_se = list()
VaR_CI = list()
ES_CI = list()

B = 1000
nsamp = nrow(data)
# VaR_samp = matrix(NA, B, ncol(data) - 2)
# ES_samp = matrix(NA, B, ncol(data) - 2)

for (i in 1:(ncol(data) - 2)) {
    VaR_samp = rep(NA, B)
    ES_samp = rep(NA, B)
    for (j in 1:B) {
        ret_samp = sample(data[, i + 2], size = nsamp, replace = T)
        VaR_samp[j] = VaR(100000, na.omit(ret_samp), 0.05)
        ES_samp[j] = ES(100000, na.omit(ret_samp), 0.05)
    }
    name = stock_names[i]
    VaR_se[[name]] = sd(VaR_samp)
    ES_se[[name]] = sd(ES_samp)
    VaR_CI[[name]][1] = quantile(VaR_samp, .25)
    VaR_CI[[name]][2] = quantile(VaR_samp, .975)
    ES_CI[[name]][1] = quantile(ES_samp, .25)
    ES_CI[[name]][2] = quantile(ES_samp, .975)
}
```

```{r}
# se and 95% CI for VaR, ES of portfolios using boostrap
VaR_p_se = list()
ES_p_se = list()
VaR_p_CI = list()
ES_p_CI = list()

B = 1000
nsamp_p = nrow(ret_p)
# VaR_samp = matrix(NA, B, ncol(data) - 2)
# ES_samp = matrix(NA, B, ncol(data) - 2)

for (i in 1:(ncol(ret_p))) {
    VaR_p_samp = rep(NA, B)
    ES_p_samp = rep(NA, B)
    for (j in 1:B) {
        ret_p_samp = sample(ret_p[, i], size = nsamp_p, replace = T)
        VaR_p_samp[j] = VaR(100000, na.omit(ret_p_samp), 0.05)
        ES_p_samp[j] = ES(100000, na.omit(ret_p_samp), 0.05)
    }
    name = colnames(ret_p)
    VaR_p_se[[name]] = sd(VaR_p_samp)
    ES_p_se[[name]] = sd(ES_p_samp)
    VaR_p_CI[[name]][1] = quantile(VaRv_samp, .25)
    VaR_p_CI[[name]][2] = quantile(VaRv_samp, .975)
    ES_p_CI[[name]][1] = quantile(ES_p_samp, .25)
    ES_p_CI[[name]][2] = quantile(ES_p_samp, .975)
}
```


###### Copula
```{r}
# set up 
df = matrix(NA, nrow(data), ncol(data) - 2)

for (i in 1:(ncol(data) - 2)) {
    df[, i] = rank(data[, i + 2]) / (nrow(data) + 1)
}
```

```{r}
# copulas for assets
library(copula)

# frank copula 
cop_f = archmCopula(family = "frank", dim = 15, param = 2)
ft_f = fitCopula(copula = cop_f, data = df, method = "ml")
aic_f = AIC(ft_f)
bic_f = BIC(ft_f)

clayton copula
cop_cl = archmCopula(family = "clayton", dim = 15, param = 2)
ft_cl = fitCopula(copula = cop_cl, data = df, method = "ml")
aic_cl = AIC(ft_cl)
bic_cl = BIC(ft_cl)

# gumbel copula
cop_g = archmCopula(family = "gumbel", dim = 15, param = 2)
ft_g = fitCopula(copula = cop_g, data = df, method = "ml")
aic_g = AIC(ft_g)
bic_g = BIC(ft_g)

# joe copula
cop_j = archmCopula(family = "joe", dim = 15, param = 2)
ft_j = fitCopula(copula = cop_j, data = df, method = "ml")
aic_j = AIC(ft_j)
bic_j = BIC(ft_j)

# t copula
cop_t = tCopula(param = rep(0, 105), dim = 15, dispstr = "un", df = 15)
ft_t = fitCopula(copula = cop_t, data = df, method = "ml")
aic_t = AIC(ft_t)
bic_t = BIC(ft_t)

# normal copula
cop_n = normalCopula(param = rep(0, 105), dim = 15, dispstr = "un")
ft_n = fitCopula(copula = cop_n, data = df, method = "ml")
aic_n = AIC(ft_n)
bic_n = BIC(ft_n)

# summary
data.frame(Copula = c("frank", "clayton", "gumbel", "joe", "t", "normal"), 
AIC = c(aic_f, aic_cl, aic_g, aic_j, aic_t, aic_n), BIC = c(bic_f, bic_cl, bic_g, bic_j, bic_t, bic_n))
```

The t copula fits the best since it has the lowest AIC and BIC. 
 Copula       AIC       BIC
1   frank -281.1542 -278.2790
2 clayton -370.9711 -368.0959
3  gumbel -320.2901 -317.4149
4     joe -212.8627 -209.9875
5       t -712.9468 -408.1759
6  normal -703.7758 -401.8801